Weighted Model Counting with Conditional Weights

### 2020-08-19 paper/paper.tex
Describe experimental results

\section{Related Work}

@@

In this section, we describe an experimental study comparing the five WMC
encodings for Bayesian networks when run with ADDMC. The experiments were run on
Intel Xeon Gold 6138 and Intel Xeon E5-2630 processors\footnote{Each instance is
run on the same processor for all encodings.} with an \SI{8}{\gibi\byte} memory
limit.

For each Bayesian network, we need to choose a probability to compute. Whenever
a Bayesian network comes with an evidence file, we compute the probability of
evidence. Otherwise, let $X$ denote the last-mentioned vertex in the Bayesian
network. If $\mathsf{true}$ is a valid value of $X$, we compute the marginal
probability of $X = \mathsf{true}$. Otherwise, we pick the value of $X$ which is
listed first and calculate its marginal probability.

@@

\paragraph{Data.} For experimental data, we use the Bayesian networks available
with Ace and
Cachet\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}, most of
which happen to be binary. We classify them into the following seven categories:
\begin{itemize*}
\item DQMR and
\item Grid networks as described by \citet{DBLP:conf/aaai/SangBK05},
\item Mastermind, and
\item Random Blocks from the work of \citet{DBLP:journals/ijar/ChaviraDJ06},
\item remaining binary Bayesian networks that include Plan Recognition
  \cite{DBLP:conf/aaai/SangBK05}, Friends and Smokers, Students and Professors
  \cite{DBLP:journals/ijar/ChaviraDJ06}, and \texttt{tcc4f}, and
\item non-binary classic Bayesian networks (\texttt{alarm}, \texttt{diabetes},
  \texttt{hailfinder}, \texttt{mildew}, \texttt{munin1}--\texttt{4},
  \texttt{pathfinder}, \texttt{pigs}, \texttt{water}).
\end{itemize*}
% TODO: make sure to talk about the table (and mention that cd06 has its niche)

\begin{figure}
  \centering
  \input{cumulative.tex}%
  \caption{Cumulative number of instances solved by ADDMC over time using each
    encoding}
\end{figure}

@@

% TODO: check if the total checks out

\begin{table}
  \centering
  \caption{The numbers of instances (out of 1466) solved by ADDMC with each
    encoding (uniquely, faster than with others, and in total) under the
    described time and memory constraints}%
  \begin{tabular}{lrrr}
    \toprule
    Encoding & Unique & Fastest & Total \\
    \midrule
    \texttt{cd05} & 2 & 3 & 380 \\
    \texttt{cd06} & 0 & 0 & 363 \\
    \texttt{cw} & 234 & 935 & 980 \\
    \texttt{d02} & 40 & 111 & 797 \\
    \texttt{sbk05} & 12 & 44 & 765 \\
    \bottomrule
  \end{tabular}
\end{table}
% TODO: another observation: cw was the fastest in 935 out of the 980 instances
% that it solved (95.4%)

\begin{figure*}
  \centering
  \input{scatter.tex}%
  \caption{ADDMC inference time using \texttt{cw} compared to \texttt{d02}
    (left) and \texttt{sbk05} (right) for each instance in all data sets.}
\end{figure*}

% TODO: write. then run through a grammar checker.
\paragraph{Observations.}
\begin{itemize}
\item The order of the encodings from best to worst is exactly the opposite of
  that in the previous literature.
\item Our encoding is particularly promising on instances of Grid networks.
  They are set up so that the entire Bayesian network is relevant to the
  query.
\item However, \texttt{cw} struggles with instances from Mastermind and
  Random Blocks domains (but so does \texttt{sbk05}). We conjecture that this
  is so either because the particular structure of these problems confuse the
  heuristics used by ADDMC or because these instances allow for significant
  simplifications that are exploited by \texttt{d02} but not \texttt{cw}.
\item An observation from the cumulative plot: \texttt{cw} solves the same
  number of instances in \SI{8.144}{\second} as \texttt{d02} does in
  \SI{1000}{\second}.
\end{itemize}

@@

\paragraph{Explaining the performance benefits.} Let $n = |\mathcal{V}|$ be the
number of vertices in the Bayesian network, $d = \max_{X \in \mathcal{V}}
|\mathrm{pa}(X)|$ the maximum in-degree (i.e., number of parents), and $v =
\max_{X \in \mathcal{V}} |\im X|$ the maximum number of values per variable.
\Cref{tbl:asymptotes} shows how \texttt{cw} has both fewer variables and fewer
ADDs than any other encoding. However, note that these are upper bounds and most
encodings (including \texttt{cw}) can be smaller in certain situations (e.g.,
with binary random variables or when a CPT has repeating probabilities). We
equate clauses and ADDs (more specifically, factors of the function $\phi$ from
\cref{alg:encoding}) here because ADDMC interprets each clause of any WMC
encoding as a multiplicative factor of the ADD that represents the entire WMC
instance \cite{DBLP:conf/aaai/DudekPV20}. For literal-weight encodings, each
weight is also a factor, but that does not affect the asymptotic bounds in the
table.

>>>>

\section{Related Work} \label{sec:related}

@@

In this section, we describe an experimental study comparing the five WMC
encodings for Bayesian networks when run with ADDMC. The experiments were run on
servers with Intel Xeon Gold 6138 and Intel Xeon E5-2630
processors\footnote{Each instance is run on the same processor for all
  encodings.} with an \SI{8}{\gibi\byte} memory limit and a \SI{1000}{\second}
timeout.

For each Bayesian network, we need to choose a probability to compute. Whenever
a Bayesian network comes with an evidence file, we compute the probability of
evidence. Otherwise, let $X$ denote the last-mentioned vertex in the Bayesian
network. If $\mathsf{true} \in \im X$, then we compute the marginal probability
of $X = \mathsf{true}$. Otherwise, we pick the value of $X$ which is listed
first and calculate its marginal probability.

@@

For experimental data, we use the Bayesian networks available with Ace and
Cachet\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}, most of
which happen to be binary. We classify them into the following seven categories:
\begin{itemize*}
\item DQMR and
\item Grid networks as described by \citet{DBLP:conf/aaai/SangBK05},
\item Mastermind, and
\item Random Blocks from the work of \citet{DBLP:journals/ijar/ChaviraDJ06},
\item remaining binary Bayesian networks that include Plan Recognition
  \cite{DBLP:conf/aaai/SangBK05}, Friends and Smokers, Students and Professors
  \cite{DBLP:journals/ijar/ChaviraDJ06}, and \texttt{tcc4f}, and
\item non-binary classic Bayesian networks (\texttt{alarm}, \texttt{diabetes},
  \texttt{hailfinder}, \texttt{mildew}, \texttt{munin1}--\texttt{4},
  \texttt{pathfinder}, \texttt{pigs}, \texttt{water}).
\end{itemize*}

\begin{figure}
  \centering
  \input{cumulative.tex}%
  \caption{Cumulative number of instances solved by ADDMC over time using each
    encoding}
  \label{fig:cumulative}
\end{figure}

@@

\begin{table}
  \centering
  \caption{The numbers of instances (out of 1466) solved by ADDMC with each
    encoding (uniquely, faster than with others, and in total) under the
    described constraints}%
  \begin{tabular}{lrrr}
    \toprule
    Encoding & Unique & Fastest & Total \\
    \midrule
    \texttt{cd05} & 2 & 3 & 380 \\
    \texttt{cd06} & 0 & 0 & 363 \\
    \texttt{cw} & 234 & 935 & 980 \\
    \texttt{d02} & 40 & 111 & 797 \\
    \texttt{sbk05} & 12 & 44 & 765 \\
    \bottomrule
  \end{tabular}
  \label{tbl:tallies}
\end{table}

\begin{figure*}
  \centering
  \input{scatter.tex}%
  \caption{ADDMC inference time using \texttt{cw} compared to \texttt{d02}
    (left) and \texttt{sbk05} (right) for each instance in all data sets.}
  \label{fig:scatter}
\end{figure*}

According to the cumulative plot in \cref{fig:cumulative}, \texttt{cw} is
consistently able to solve more instances with any given time limit than any
other encoding, while the order of the remaining encodings from best to worst is
exactly the opposite of that in previous literature (cf. \cref{sec:related}).
Specifically, \texttt{cw} solves the same number of instances in
\SI{8.144}{\second} as \texttt{d02} does in \SI{1000}{\second}, which is the
second-best encoding in our experiments. Furthermore, \cref{tbl:tallies} shows
that \texttt{cw} was the fastest in 935 out of the 980 (i.e.,
\SI{95.4}{\percent}) instances that it managed to solve. Finally, according to
the scatter plots in \cref{fig:scatter}, \texttt{cw} is particularly promising
on Grid networks while losing to \texttt{d02} on some instances of the
Mastermind data set. We conjecture that this is so either because the particular
structure of the problem confuses the heuristics used by ADDMC or because these
instances allow for significant simplifications that are exploited by
\texttt{d02} but not \texttt{cw}.

@@

\paragraph{Explaining the performance benefits.} Let $n = |\mathcal{V}|$ be the
number of vertices in the Bayesian network, $d = \max_{X \in \mathcal{V}}
|\mathrm{pa}(X)|$ the maximum in-degree (i.e., the number of parents), and $v =
\max_{X \in \mathcal{V}} |\im X|$ the maximum number of values per variable.
\Cref{tbl:asymptotes} shows how \texttt{cw} has fewer variables and fewer
ADDs than any other encoding. However, note that these are upper bounds and most
encodings (including \texttt{cw}) can be smaller in certain situations (e.g.,
with binary random variables or when a CPT has repeating probabilities). We
equate clauses and ADDs (more specifically, factors of the function $\phi$ from
\cref{alg:encoding}) here because ADDMC interprets each clause of any WMC
encoding as a multiplicative factor of the ADD that represents the entire WMC
instance \cite{DBLP:conf/aaai/DudekPV20}. For literal-weight encodings, each
weight is also a factor, but that does not affect the asymptotic bounds in the
table.
